Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-09-05 write_up/mixfishsim_ecolmod.tex
Trimming the fat

\begin{frontmatter}
\title{\emph{MixFishSim}: highly resolved spatiotemporal simulations for
	exploring mixed fishery dynamics}

@@

We simulated 50 years of fishing and used the results to draw inference on the
underlying community structures. We compared inference from commercial catch to
a simulated fixed-site sampling design and the true underlying community. We i)
establish the potential limitations of fishery-dependent data in providing a
robust picture of spatiotemporal distributions; and then ii) simulated an area
closure based on areas defined from the different data sources at a range of
temporal and spatial resolutions to assess their effectiveness on reducing
catches of a fish population. \\

@@

\replaced[id=PD]{Adaptive spatial management strategies have}{Use of spatial
	management as a tools} been proposed as a way of reducing overquota
discards \added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.
\deleted[id=PD]{its} However, implementation is restricted by lack of knowledge
of fish and fishery spatiotemporal dynamics and understanding of the scale at
which processes become important for management. Understanding the correct
scale for spatial measures is crucial for implementing effective solutions
which  minimising economic impact \citep{Dunn2016}. For example, the problem
can be to identify a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species.\\

@@

\replaced[id=PD]{A}{We developed and implemented a fully spatially explicit
	simulation model with a} \replaced{simulation model that is modular and
	discrete-event based was developed. This approach enables efficient
	computation by allowing for}{approach, where}
\added[id=PD]{sub-}modules \deleted[id=PD]{are} implemented on time-scales
appropriate to capture the characteristic of the \replaced[id=PD]{different
	processes}{process modelled} (Figure 1).  \added[id=PD]{The following
	sub-modules were included to capture the full system: 1) Population
	dynamics, 2) Recruitment dynamics, 3) Population movement, 4) fishery
	dynamics.}\\

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics (including fishing and natural
mortality which are instantaneous rates, and growth of the population biomass)
operate on a daily time-step\replaced[id=PD]{, while p}{. P}opulation movement
occurs on a weekly time-step\replaced[id=PD]{. R}{, while r}ecruitment
\replaced[id=PD]{takes place}{occurs} periodically each year for a set time
\replaced[id=PD]{duration}{period (e.g. 3 weeks)} specified for each
population\deleted[id=PD]{.}\added[id=PD]{, while the fishing module operates
	on a tow-by-tow basis (i.e. multiple events a day)}.
\deleted[id=PD]{Here we describe each of the model components; 1) Population
	dynamics, 2) Recruitment dynamics, 3) Population movement dynamics, 4)
	fishery dynamics.}

@@

Recruitment is modelled through a function relating the adult biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \cdot S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot \mathrm{e}^{(\alpha - \beta \cdot B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In our example
application the Beverton-Holt form of stock recruit relationship was used for
all populations though either functional form can be chosen.

@@

\deleted[id=PD]{the habitat}We first define\added[id=PD]{d} a Gaussian random
field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian with a \textit{Matérn}
\deleted[id=PD]{family of}covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{where}{one where} the correlation strength weakens with
distance\deleted[id=PD]{(i.e. the correlation between $S(x)$ and $S(x')$
	decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{This enables us to model the spatial autocorrelation observed in
	animal populations where density is more similar in nearby locations
	\citep{Tobler1970, F.Dormann2007}} and we change the parameters to
implement different spatial structures for the populations. \\

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, that simulates a Gaussian Random Field process given a
user defined error model and correlation structure. We define a stationary
habitat field and combine with a temporally dynamic thermal tolerance field to
imitate two key drivers of population dynamics. Each population
\replaced[id=PD]{was}{is} initialised at a single location, and subsequently
move\replaced[id=PD]{d}{s} according to a probabilistic
distrib\added[id=PD]{u}tion based on habitat suitability (represented by the
normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{\mathrm{e}^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below).\\

@@

\added[id=PD]{The final process results in a population structure and movement
	pattern unique to each species, with population movement occurring on a
	weekly basis. The decision to model population movement on a weekly
	timescale was to reflect that fish tend to aggregate in species
	specific locations that have been observed to last around one to two
	weeks \citep{Poos2007}.  Therefore this process approximated the
	demographic shifts in fish populations throughout a year with seasonal
	spawning patterns (e.g.  Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular species; trip-level decisions, that determined the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determined movement from one fishing spot to another within a trip. Together,
these elements implemented an explore-exploit type strategy for individual
vessels to maximise their catch from an unknown resource distribution
\cite{Bailey2018}.  The decision to use an individual based model for fishing
vessels was taken because fishers are heterogeneous in their location choice
behaviour due to different objectives, risk preference and targeting preference
\citep{VanPutten2012a}. Therefore in the simulations fleet dynamics are the
productive of individual experiences rather than pre-defined group dynamics.

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, $Q_{fl}$, and a population specific efficiency,
${Q_{fl, p}}$.  Thus, the product of these parameters [$Q_{fl} \cdot Q_{fl,
	p}$] affect\replaced[id=PD]{s}{ed} the overall catch rates for the
fleet and the preferential targeting of one population over another. This, in
combination with the parameter choice for the step-function
\added[id=PD]{defined below} (as well as some randomness from the exploratory
fishing process) determine\replaced[id=PD]{d}{s} the preference of fishing
locations for the fleet.  All species prices \replaced[id=PD]{were}{are} kept
the same across fleets \replaced[id=PD]{and seasons}{, though can be made to
	vary seasonally}.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by \deleted[id=JJ]{relating} recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$),
\begin{equation}
Rev = \sum_{p=1}^{P} \replaced[id=PD]{L}{C}_{p} \cdot Pr_{p} 
\end{equation}
where $\replaced[id=PD]{L}{C}_{p}$ is \replaced[id=PD]{landings}{catch} of a
population $p$, and $Pr_{p}$ price of a population. Here, when fishing is
successful vessels remain in a similar location and continue to exploit the
local fishing grounds. When unsuccessful, they move some distance away from
the current fishing location.  The movement distance retains some degree of
stochasticity, that can be controlled separately, but is determined by the
relationship: 
\begin{equation}
	StepL = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		(log(\frac{\beta_{1}}{\beta_{3}}))} * Rev
\end{equation}\todo{\added[id=JJ]{So
			step length increases with increasingly gross
			revenue?}\added[id=PD]{No, the opposite}}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + StepL \cdot \cos (\frac{\pi \cdot Br}{180}), \\
             & y1 + StepL \cdot \sin (\frac{\pi \cdot Br}{180}) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $k$ the concentration parameter from the von \replaced[id=JJ]{M}{m}ises
distribution that we correlate with the revenue so that $k = (Rev + 1 /
RefRev) * max_{k}$, where $max_{k}$ is the maximum concentration value, $k$,
and $RefRev$ is parametrised as for $\beta_{3}$ in the step length function. A
realised example of the step length and turning angle relationships to revenue
can be seen at Figure \ref{S-fig:15}.

@@

We parametrised the simulation model for four populations with different
demographics; growth rates, natural mortality and recruitment functions (Table
\ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and temperature
tolerances (Figures \ref{S-fig:3}, \ref{S-fig:4}) were defined to be unique to
each population resulting in differently weekly distribution patterns (Figures
\ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the populations was assumed
to have two defined spawning areas that result in the populations moving
towards these areas in pre-defined weeks (Figure \ref{S-fig:2}) with
population-specific movement rates (Table \ref{tab:1}). In such a
configuration, the individual habitat preferences and thermal tolerances result
in different spatial habitat use for each population (Figure \ref{fig:9}) and
consequently different seasonal exploitation patterns (Fishing mortality in
Figure \ref{fig:10}).

>>>>

\begin{frontmatter}
\title{Working title: Highly resolved spatiotemporal simulations for
	exploring mixed fishery dynamics}

@@

We simulated 50 years of fishing and used the results to draw inference on the
underlying community structures. We compared inference from commercial catch to
a simulated fixed-site sampling design and the true underlying community. We i)
establish the potential limitations of fishery-dependent data in providing a
robust picture of spatiotemporal distributions; and then ii) simulated an area
closure based on areas defined from the different data sources at a range of
temporal and spatial resolutions. We assessed the effectiveness of the closures
in reducing catches of a fish population. \\

@@

\replaced[id=PD]{Adaptive spatial management strategies have}{Use of spatial
	management as a tools} been proposed as a way of reducing over-quota
discards \added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.
\deleted[id=PD]{its} However, implementation is restricted by lack of knowledge
of fish and fishery spatiotemporal dynamics and understanding of the scale at
which processes become important for management. Understanding the correct
scale for spatial measures is crucial for implementing effective solutions
which  minimising economic impact \citep{Dunn2016}. For example, the problem
can be to identify a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species.\\

@@

\replaced[id=PD]{An}{We developed and implemented a fully spatially explicit
	simulation model with a} \replaced{event-based simulation model of a
	hypothetical fishery was developed as a software package
	(`MixFishSim'). The modular approach enabled efficient computation by allowing
	for}{approach, where} \added[id=PD]{sub-}modules \deleted[id=PD]{are}
implemented on time-scales appropriate to capture the characteristic of the
\replaced[id=PD]{different processes}{process modelled} (Figure 1).
\added[id=PD]{The following sub-modules were included to capture the full
	system: 1) Population dynamics, 2) Recruitment dynamics, 3) Population
	movement, 4) fishery dynamics.}\\

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics operate on a daily
time-step\replaced[id=PD]{, while p}{. P}opulation movement occurs on a weekly
time-step, with the fishing module operating on a tow-by-tow basis (i.e.
multiple events a day).  \deleted[id=PD]{Here we describe each of the model
components; 1) Population dynamics, 2) Recruitment dynamics, 3) Population
movement dynamics, 4) fishery dynamics.}

@@

Recruitment is modelled through a function relating to adult biomass. In
\emph{MixFishSim}, it can be either as a stochastic Beverton-Holt stock-recruit
form \citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \cdot S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot \mathrm{e}^{(\alpha - \beta \cdot B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases.

@@

\deleted[id=PD]{the habitat}We first define\added[id=PD]{d} a Gaussian random
field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian with a \textit{Matérn}
\deleted[id=PD]{family of}covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{where}{one where} the correlation strength weakens with
distance\deleted[id=PD]{(i.e. the correlation between $S(x)$ and $S(x')$
	decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{This enables us to model the spatial autocorrelation observed in
	animal populations where density is more similar in nearby locations
	\citep{Tobler1970, F.Dormann2007}} and we change the parameters to
implement different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field and combine with a temporally dynamic thermal tolerance field to
imitate two key drivers of population dynamics. Each population
\replaced[id=PD]{was}{is} initialised at a single location, and subsequently
move\replaced[id=PD]{d}{s} according to a probabilistic
distrib\added[id=PD]{u}tion based on habitat suitability (represented by the
normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{\mathrm{e}^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below).\\

@@

\added[id=PD]{The final combined process results in a population structure and
	movement pattern unique to each species, with population movement
	occurring on a weekly basis. The decision to model population movement
	on a weekly timescale was to reflect that fish tend to aggregate in
	species specific locations that have been observed to last around one
	to two weeks \citep{Poos2007}. Therefore this process approximated the
	demographic shifts in fish populations throughout a year with seasonal
	spawning patterns (Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular species; trip-level decisions, that determined the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determined movement from one fishing spot to another within a trip. Together,
these elements implemented an explore-exploit type strategy for individual
vessels to maximise their catch from an unknown resource distribution
\citep{Bailey2018}. The decision to use an individual based model for fishing
vessels was taken because fishers are heterogeneous in their location choice
behaviour due to different objectives, risk preference and targeting preference
\citep{VanPutten2012a}. Therefore in the simulations fleet dynamics are the
productive of individual experiences rather than pre-defined group dynamics.

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, $Q_{fl}$, and a population specific efficiency,
${Q_{fl, p}}$.  Thus, the product of these parameters [$Q_{fl} \cdot Q_{fl,
	p}$] affect\replaced[id=PD]{s}{ed} the overall catch rates for the
fleet and the preferential targeting of one population over another. This, in
combination with the parameter choice for the step-function
\added[id=PD]{defined below} (as well as some randomness from the exploratory
fishing process) determine\replaced[id=PD]{d}{s} the preference of fishing
locations for the fleet.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by \deleted[id=JJ]{relating} recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$),
\begin{equation}
Rev = \sum_{p=1}^{P} \replaced[id=PD]{L}{C}_{p} \cdot Pr_{p} 
\end{equation}
where $\replaced[id=PD]{L}{C}_{p}$ is \replaced[id=PD]{landings}{catch} of a
population $p$, and $Pr_{p}$ price of a population. All species prices were
kept the same across fleets and seasons. Here, when fishing is successful
vessels remain in a similar location and continue to exploit the local fishing
grounds. When unsuccessful, they move some distance away from the current
fishing location.  The movement distance retains some degree of stochasticity,
that can be controlled separately, but is determined by the
relationship: 
\begin{equation}
	StepL = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		(log(\frac{\beta_{1}}{\beta_{3}}))} * Rev
\end{equation}\todo{\added[id=JJ]{So
			step length increases with increasingly gross
			revenue?}\added[id=PD]{No, the opposite}}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + StepL \cdot \cos (\frac{\pi \cdot Br}{180}), \\
             & y1 + StepL \cdot \sin (\frac{\pi \cdot Br}{180}) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $k$ the concentration parameter from the von \replaced[id=JJ]{M}{m}ises
distribution that we correlate with the revenue so that $k = (Rev + 1 /
RefRev) * max_{k}$, where $max_{k}$ is the maximum concentration value, $k$,
and $RefRev$ is parametrised as for $\beta_{3}$ in the step length function. A
realised example of the step length and turning angle relationships to revenue
can be seen at Figure \ref{S-fig:15}.

@@

We parametrised the simulation model for four populations with different
demographics; growth rates, natural mortality and recruitment parameters (Table
\ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and temperature
tolerances (Figures \ref{S-fig:3}, \ref{S-fig:4}) were defined to be unique to
each population resulting in differently weekly distribution patterns (Figures
\ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the populations was assumed
to have two defined spawning areas that result in the populations moving
towards these areas in pre-defined weeks (Figure \ref{S-fig:2}) with
population-specific movement rates (Table \ref{tab:1}). In such a
configuration, the individual habitat preferences and thermal tolerances result
in different spatial habitat use for each population (Figure \ref{fig:9}) and
consequently different seasonal exploitation patterns (Fishing mortality in
Figure \ref{fig:10}).
